---
title: "Calculate LAI"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(treetemp)
library(devtools)
load_all()
library(ggplot2)
library(dplyr)
library(stringr)
library(directlabels)

# Dir for calculated lai
#dir_lai<-"D:/MetOff_data/fieldwork/LAI_results"
dir_lai<-"/Users/jonathanmosedale/Library/CloudStorage/OneDrive-UniversityofExeter/MetOff_data/field_data/LAI_results"

# Dir containing all .dat files (can be in subdirectories)
dir_picroot<-"D:/MetOff_data/fieldwork/Shared Canopy Pic Trial"
pic_dirs<-list.dirs(dir_picroot,recursive=FALSE)

# dir_data<-file.path("D:/MetOff_data/fieldwork/Shared Canopy Pic Trial",pic_dirs[1])


# Set type of LAI calculation to usee - see read_lai_results
#lai<-'LiCor' 

```


# Get LAI data for each site
Processes .dat files written from hemisfer software. Extracts relevant information and LAI measurements.

Creates long data format and saves


```{r get_hemisfer_info}
# Results df for all sites
result_col_names<-c('site','spp','tree','treeid','date','lai000n','lai100n','lai000s','lai100s','lai0','lai100')

allresults_df<-data.frame(matrix(ncol = length(result_col_names), nrow = 0))
colnames(allresults_df)<-result_col_names
result_col_names<-c("filepath","filename","date","treeid","site","spp","tree","asp","height",  
 "lai","angle","transm","skyview","lggaps","Fmv","clump1","clump2","clump3", "clump")
allresults_df<-data.frame(matrix(ncol = length(result_col_names), nrow = 0))
colnames(allresults_df)<-result_col_names

  
for(dir_data in pic_dirs){
  print(dir_data)

  # Results for this site
  results_df<-data.frame(matrix(ncol = length(result_col_names), nrow = 0))
  colnames(results_df)<-result_col_names

  # Get lai results and metadata of files
  files<-list.files(dir_data,pattern="*.dat",full.names = TRUE, recursive=TRUE) # recursive - includes all subdirs
  files<-files[which(basename(files)!='LAI.dat')]
  fdata<-data.frame(t(sapply(files,read_lai_results)))
  fdata<-as.data.frame(lapply(fdata,unlist),row.names = NULL)
  fdata<-fdata[order(fdata$date,fdata$treeid,fdata$height,fdata$asp),]

  # Reorganise data -  Date, TreeID n/s, Lai100, Lai000 - CORRECT ADJACENT DATES!!
  for(tree in unique(fdata$treeid)){
    treedata<-fdata[which(fdata$treeid==tree),]
    
    # Check and correct if dates within 1 day
    tdates<-unique(as.POSIXlt(treedata$date))
    tdates<-tdates[order(tdates)]
    for(n in 2:length(tdates)){
      #print(as.numeric(tdates[n]-tdates[n-1]))
      if (as.numeric(tdates[n]-tdates[n-1])<=5){ 
        print(paste('Correcting dates for tree',tree))
        treedata$date[which(treedata$date==tdates[n-1])] <- as.character(tdates[n])
      }
    }
    results_df<-rbind(results_df,treedata)
  }
  allresults_df<-rbind(allresults_df,results_df)
} # dir_data

write.csv(allresults_df,file.path(dir_lai,paste('lai_allsites_results.csv',sep='_')))
```
## Missing data
2024-05-08 be_oak_1 N000 reading
2024-06-20 cw_oak_3 N100 reading
2024-06-19 wh_oak_2 S000 reading 
2024-04-29 wh_oak_5 N100 reading

## Estimate of clumping
High variability between sites and trees but nevertheless NSP generally lower and more variable values than oak.


Oak mean clunp = 0.95 (lower variability apart from AH and one tree in CW )
NSP mean clump = 0.89 (high between tree variability (partic at WH least at AH)

Omega = LAIe/LAI
small omega - ratio needl and twig  - 


To use tree average values across time period

```{r plot_clumping}
allresults_df<-read.csv(file.path(dir_lai,paste('lai_allsites_results.csv',sep='_')))

# Mean clump (from rings 1 to 3) ro each height and aspect
clumpwide_df<-as.data.frame(allresults_df %>% 
                tidyr::pivot_wider(id_cols=c(treeid,date,site,spp,tree),
                            names_from=c(asp,height),
                            names_glue = "{asp}_{height}_clump", 
                            values_from=c(clump) ) )
clumpwide_df$clump_0<-apply(as.matrix(clumpwide_df[,6:7]),1,mean,na.rm=T)
clumpwide_df$clump_100<-apply(as.matrix(clumpwide_df[,8:9]),1,mean,na.rm=T)

# Plot clumping by sapp, site
height<-100
figlist<-list()
for (site in unique(clumpwide_df$site)){
  figlist[[site]]<-plot_param(clumpwide_df,site,height,'clump')
  figname<-paste('clump',site,height,"treeplots.png",sep='_')
  ggsave(file.path(dir_lai,figname),figlist[[site]])
}
lapply(figlist,print)

height<-0
figlist<-list()
for (site in unique(clumpwide_df$site)){
  figlist[[site]]<-plot_param(clumpwide_df,site,height,'clump')
  figname<-paste('clump',height,site,"treeplots.png",sep='_')
  ggsave(file.path(dir_lai,figname),figlist[[site]])
}
lapply(figlist,print)

clump_stats<-clumpwide_df %>% group_by(spp, site) %>% 
  summarise( mean100 = mean(clump_100,na.rm=T), sd100 = sd(clump_100,na.rm=T),mean0=mean(clump_0,na.rm=T),sd0=sd(clump_0,na.rm=T))

print(clump_stats)

spp_clump<-clumpwide_df %>% group_by(spp) %>% 
  summarise( mean100 = mean(clump_100,na.rm=T), sd100 = sd(clump_100,na.rm=T),mean0=mean(clump_0,na.rm=T),sd0=sd(clump_0,na.rm=T))

print(spp_clump)

saveRDS(clumpwide_df,file.path(dir_lai,paste('clumping_allsites.Rds',sep='_')))
```


### Angle estimates
Very high variation across time and trees - oak included many 0 values (completely flat). 
Could be calculated per site to get more robust estimate.
LAI-2000 LAI is independent of leaf angle
In Thimonier LAI is dependent on leaf angle - angle could be corrected to exclude extreme values (~0 or)

Mean values are:
NSP = 43
Oak = 18 (incl 0 values) or 36 (excluding 0 values)

alpha leaf inclination

```{r angle_plots}
# Angle - very variable!
anglewide_df<-as.data.frame(allresults_df %>% 
                tidyr::pivot_wider(id_cols=c(treeid,date,site,spp,tree),
                            names_from=c(asp,height),
                            names_glue = "{asp}_{height}_angle", 
                            values_from=c(angle) ) )
anglewide_df$angle_0<-apply(as.matrix(anglewide_df[,6:7]),1,mean,na.rm=T)
anglewide_df$angle_100<-apply(as.matrix(anglewide_df[,8:9]),1,mean,na.rm=T)

height<-100
figlist<-list()
for (site in unique(anglewide_df$site)){
  figlist[[site]]<-plot_param(anglewide_df,site,height,'angle')
  figname<-paste('angle_',site,height,"treeplots.png",sep='_')
  ggsave(file.path(dir_lai,figname),figlist[[site]])
}
lapply(figlist,print)

height<-0
figlist<-list()
for (site in unique(anglewide_df$site)){
  figlist[[site]]<-plot_param(anglewide_df,site,height,'angle')
  figname<-paste('angle_',site,height,"treeplots.png",sep='_')
  ggsave(file.path(dir_lai,figname),figlist[[site]])
}
lapply(figlist,print)


# anglewide_df$angle_100<-ifelse(anglewide_df$angle_100==0,NA,anglewide_df$angle_100)
# anglewide_df$angle_0<-ifelse(anglewide_df$angle_0==0,NA,anglewide_df$angle_0)

sppste_angle<-anglewide_df %>% group_by(spp, site) %>% 
  summarise( mean100 = mean(angle_100,na.rm=T), sd100 = sd(angle_100,na.rm=T),mean0=mean(angle_0,na.rm=T),sd0=sd(angle_0,na.rm=T))

print(sppste_angle)

spp_angle<-anglewide_df %>% group_by(spp) %>% 
  summarise( mean100 = mean(angle_100,na.rm=T), sd100 = sd(angle_100,na.rm=T),mean0=mean(angle_0,na.rm=T),sd0=sd(angle_0,na.rm=T))

print(spp_angle)

saveRDS(anglewide_df,file.path(dir_lai,paste('leafangles_allsites.Rds',sep='_')))

```


## Plot tree LAIs
Similar range in LAI values (0.4 to 4.5) for each spp but for Norway spruce such variation is primarily between trees and for oak over time.

Choice of using:
LiCaor-2000 method - LAI calc independent of estimate of angle 
  LAI (uncorrected) NSP 0.81 to 3.65
  LAI (uncorrected) Oak 0.42 to 4.48

Thimonier LAI dependent on calc of angle
  LAI (uncorrected) NSP 0.78 to 4.12
  LAI (uncorrected) Oak 0.36 to 4.34


```{r plot}
library(ggplot2)
library(directlabels)
allresults_df<-read.csv(file.path(dir_lai,paste('lai_allsites_results.csv',sep='_')))

# What LAI to plot?
height<-'100'
#height<-'0'
lai<-"lc"
lai<-"t"

laiwide_df<-as.data.frame(allresults_df %>% 
                pivot_wider(id_cols=c(treeid,date,site,spp,tree),
                            names_from=c(asp,height),
                            names_glue = "{asp}_{height}_lai", # change this for lai type
                            values_from=paste0('lai_',lai) ) )     

if(height=='000') laiwide_df$lai<-apply(laiwide_df[,c(6,7)],1,mean,na.rm=TRUE)
if(height=='100') laiwide_df$lai<-apply(laiwide_df[,c(8,9)],1,mean,na.rm=TRUE)

print(paste('NSp:',range(laiwide_df$lai[which(laiwide_df$spp=='nsp')])))
print(paste('Oak:',range(laiwide_df$lai[which(laiwide_df$spp=='oak')])))

# Plot
ymin<-min(laiwide_df$lai)
ymax<-max(laiwide_df$lai)

for (site in unique(laiwide_df$site)){
  plotdata<-laiwide_df[which(laiwide_df$site==site),]
  plotdata<-plotdata[order(plotdata$treeid),]
  plotdata$label<-paste0(plotdata$spp,'_',plotdata$tree)
  fig<-ggplot(data = plotdata, aes(x = factor(date), y = lai, color = spp)) + 
    ggtitle(paste("LAI",lai,"of trees from site",site)) + ylim(ymin,ymax) +
    xlab("Date") + ylab(paste(lai,'at',height,'metres')) + 
    geom_line(aes(group = treeid)) + geom_point() +
    geom_dl(aes(label=label),method=list(dl.combine("first.points","last.points"),cex=0.7))
  figname<-paste(lai,site,"treeplots.png")
  ggsave(file.path(dir_lai,figname),fig)
  print(fig)
}

# Basic stats
sppste_lai<-laiwide_df %>% group_by(spp, site) %>% 
  summarise( mean100 = mean(lai,na.rm=T), min100 = min(lai,na.rm=T),max100=max(lai,na.rm=T))

spp_lai<-laiwide_df %>% group_by(spp) %>% 
  summarise( mean100 = mean(lai,na.rm=T), min100 = min(lai,na.rm=T),max100=max(lai,na.rm=T))

print(sppste_lai)
print(spp_lai)

```

# Curve fitting to LAI
Initial exploration using linear and polynomial model fitting.

```{r curve-fitting}
lai<-'lc'

allresults_df<-read.csv(file.path(dir_lai,paste('lai_allsites_results.csv',sep='_')))

# What LAI to plot?
height<-100
#height<-0
lai<-"lc"
# Exclude KG as few results?
sites<-c('ah','be','cw','wh')

laiwide_df<-as.data.frame(allresults_df[which((allresults_df$site %in% sites)),] %>% 
                pivot_wider(id_cols=c(treeid,date,site,spp,tree),
                            names_from=c(asp,height),
                            names_glue = "{asp}_{height}_lai", # change this for lai type
                            values_from=paste0('lai_',lai) ) )     




if(height==0) laiwide_df$lai<-apply(laiwide_df[,c(6,7)],1,mean,na.rm=TRUE)
if(height==100) laiwide_df$lai<-apply(laiwide_df[,c(8,9)],1,mean,na.rm=TRUE)

for(t in unique(laiwide_df$treeid)){
  modeldata<-laiwide_df[which(laiwide_df$treeid==t),]
  modeldata$date<-as.numeric(strftime(as.POSIXct(modeldata$date), format = "%j"))
  
  # fit pollynomial regression models up to degree 5 
  linear_model1 <- lm(lai~date, data=modeldata) 
  linear_model2 <- lm(lai~poly(date,2,raw=TRUE), data=modeldata) 
  linear_model3 <- lm(lai~poly(date,3,raw=TRUE), data=modeldata) 
  linear_model4 <- lm(lai~poly(date,4,raw=TRUE), data=modeldata) 
  linear_model5 <- lm(lai~poly(date,5,raw=TRUE), data=modeldata) 
    
  # create a basic scatterplot  
  
  plot(modeldata$date, modeldata$lai,main=t,ylim=c(0.4,4.5)) 
    
  # define x-axis values 
  x_axis <- seq(50,300,10)
    
  if(modeldata$spp[1] =="nsp"){
    lines(x_axis, predict(linear_model1, data.frame(date=x_axis)), col='green') 
    lines(x_axis, predict(linear_model2, data.frame(date=x_axis)), col='red') 
    #nsp_list[[t]]<-linear_model2
  } else{
      lines(x_axis, predict(linear_model2, data.frame(date=x_axis)), col='red') 
      lines(x_axis, predict(linear_model3, data.frame(date=x_axis)), col='purple') 
  }
}
```

### Model calculation for predicting LAI
DoY range for which there may be lai data = 69 to 269

For Norway Spruce a quadratic equation was found to fit the range of observations (although in many cases this differed little from a simple linear model fit)

For Oak a Weibull model (type 2) was found to fit observations - with the assumption that no significant leaf loss begins before the final LAI observation in late September or early October.

R pacakage ref: https://www.statforbiology.com/2020/stat_nls_usefulfunctions/#weibull-function-type-1 

Papers re shape of seasonal lai curve: https://www.mdpi.com/2072-4292/11/6/689 
https://www.researchgate.net/figure/The-seasonal-course-of-leaf-area-index-LAI-values-of-a-lysimeter-grown-Vitis-vinifera_fig2_225666686
https://bg.copernicus.org/articles/15/6885/2018/#&gid=1&pid=1

deciduous forest: https://www.researchgate.net/publication/10703196_Leaf_dynamics_of_a_deciduous_forest_canopy_No_response_to_elevated_CO_2 
Found photo derived LAI ~2 during growing season with peak of ~3 (whereas actual measured was >4)

be nsp - very littl evariation - some eg nsp2 lower LAI in summer
weibull cannot be fitted to some nsp

```{r lai_model_calc}
library(drc)
library(drcSeedGerm)

lai_models<-list()
for(t in unique(laiwide_df$treeid)){
  modeldata<-laiwide_df[which(laiwide_df$treeid==t),]
  modeldata$date<-as.numeric(strftime(as.POSIXct(modeldata$date), format = "%j"))
  n<-nrow(modeldata)
  mindoy<-min(modeldata$date)
  maxdoy<-max(modeldata$date)
  
  if(modeldata$spp[1] =="nsp"){
    model<-lm(lai~poly(date,2,raw=TRUE), data=modeldata)  
    # if likely u-shape (first and last obs > all others) then use linear model
    if(all(modeldata$lai[1]>(modeldata$lai[2:(n-1)])) &
       all(modeldata$lai[n]>(modeldata$lai[2:(n-1)]))){
      model<-lm(lai~poly(date,1,raw=TRUE), data=modeldata)} 
    # if -ve slope at beginning and +ve at end change to straight line
    testdat<-predict(model, data.frame(date=seq(mindoy,maxdoy,1)))
    if(testdat[2]-testdat[1]<0 & testdat[length(testdat)]-testdat[(length(testdat)-1)]>0) model<-lm(lai~date, data=modeldata)
    
    
  }
  if(modeldata$spp[[1]]=='oak') model <- drm(lai~date, fct =  W2.4(), data=modeldata)
  
  x_axis <- seq(60,270,10)
  plot(modeldata$date, modeldata$lai,main=t,ylim=c(0,5),xlab = "Day of Year")
  lines(x_axis, predict(model, data.frame(date=x_axis)), col='red') 

  lai_models[[t]]<-model
}

saveRDS(lai_models,file.path(dir_lai,paste0(lai,'_lai_models.Rds')))
```


### Plot LAI curves for each tree by site


```{r lai_bmodel_calc}
library(drc)
library(drcSeedGerm)
library(aomisc)
allresults_df<-read.csv(file.path(dir_lai,'lai_allsites_results.csv'),sep=',')

# try cobs
mx<-max(modeldata$lai)
mn<-min(modeldata$lai)
#con<-rbind(c(-1,max(modeldata$date),mx+0.1))
r1<-c(-1,1,mn-0.5)
r2<-c(-1,366,mn-0.5)

r1<-c(0,60,mn)
r2<-c(0,300,mn)


con<-rbind(r1,r2)
model<-cobs(modeldata$date,modeldata$lai,constraint="concave", pointwise=con)
x_axis <- seq(1,366,10)
plot(modeldata$lai~modeldata$date,main=t,xlim=c(60,280),ylim=c(0,5),xlab = "Day of Year")
lines(predict(model,x_axis),col='red')



lai_models<-list()
for(t in unique(laiwide_df$treeid)){
  modeldata<-laiwide_df[which(laiwide_df$treeid==t),]
  modeldata$date<-as.numeric(strftime(as.POSIXct(modeldata$date), format = "%j"))
  n<-nrow(modeldata)
  
  model <- drm(lai~date, fct =  DRC.bragg.3(), data=modeldata)
  

  
  x_axis <- seq(60,270,10)
  plot(modeldata$date, modeldata$lai,main=t,ylim=c(0,5),xlab = "Day of Year")
  lines(x_axis, predict(model, data.frame(date=x_axis)), col='red') 

  lai_models[[t]]<-model
}

#saveRDS(lai_models,file.path(dir_lai,paste0(lai,'_lai_models.Rds')))


lai_models<-list()
for(t in unique(laiwide_df$treeid)){
  modeldata<-laiwide_df[which(laiwide_df$treeid==t),]
  modeldata$date<-as.numeric(strftime(as.POSIXct(modeldata$date), format = "%j"))
  n<-nrow(modeldata)
  
  # if(modeldata$spp[1]=='nsp') model<-lm(lai~poly(date,2,raw=TRUE), data=modeldata) 
  if(modeldata$spp[1]=='nsp') model<-cobs(modeldata$date,modeldata$lai, constraint="concave")
  if(modeldata$spp[1]=='oak') model <- drm(lai~date, fct =  W2.4(), data=modeldata)
  
  x_axis <- seq(60,270,10)
  plot(modeldata$date, modeldata$lai,main=t,ylim=c(0,5),xlab = "Day of Year")
  if(modeldata$spp[1]=='oak') lines(x_axis, predict(model, data.frame(date=x_axis)), col='red') 
  if(modeldata$spp[1]=='nsp') lines(predict(model,x_axis))
  lai_models[[t]]<-model
}

```



### Plot model curves for all field study trees
Predicted LAI development from fitting models to observed LAI derived from canopy photos.

```{r lai_data}

# Plot example LAI curves for AH
par(mfrow=c(2,2))
layout(matrix(c(1,2,3,4), 2, 2, byrow = TRUE), widths = lcm(5), heights =lcm(5))
for(site in c('AH','BE','CW','WH')){
  trees<-trees_sf[which(trees_sf$Site==site),]
  for(t in 1:nrow(trees)){
    laimod<-lai_models[[trees$id[t]]]
    #tme<-unique(obs_df$datetime)
    tme<-unique(loggers_df$datetime)
    doy_seq<-unique(as.numeric(strftime(as.POSIXct(tme), format = "%j")))
    lai_seq<-predict(laimod, data.frame(date=doy_seq))
    plotdata<-data.frame("doy"=doy_seq,"lai"=lai_seq)
    if (trees$Tree_species[t]=='nsp') col<-'green'
    if (trees$Tree_species[t]=='oak') col<-'purple'
    if(t==1) print(plot(plotdata$doy, plotdata$lai,col=col,type='l',ylab="Model LAI", ylim=c(0,5),xlab = "Day of Year")) else lines(plotdata$doy, plotdata$lai,col=col) 
  }
}
```
