---
title: "Data Processing"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
prj <- system.file("proj", package = "terra")[1]
Sys.setenv("PROJ_LIB" = prj)
# 

# install_github("ilyamaclean/mesoclim")
# install_github("jrmosedale/mesoclimAddTrees")

library(devtools)
load_all() # Assuming run under treetemp R project/package

#library(treetemp)
library(stringr)
library(ggplot2)
library(scales)
library(tidyr)
library(mesoclim)
library(mesoclimAddTrees)
library(leaflet)
library(sf)
library(ncdf4)
library(terra)
library(dplyr)
library(dygraphs)
library(xts)
library(lubridate)

dir_root<-"D:/MetOff_data/fieldwork"
dir_tinytag<-file.path(dir_root,"Tiny Tag Data","tt_last_csv")
dir_surveytag<-file.path(dir_root,"Survey Tag Data")

dir_od<-"C:/Users/jm622/OneDrive - University of Exeter"
#dir_od<-"/Users/jonathanmosedale/Library/CloudStorage/OneDrive-UniversityofExeter"

dir_lai<-file.path(dir_od,"MetOff_data","field_data","LAI_results")

dir_out<-file.path(dir_od,"MetOff_data/field_data/Logger_results")

dir_field<-file.path(dir_od,"MetOff_data/field_data")
dir_dtm<-file.path(dir_od,"Data/Terrain50")
dir_era5<-file.path(dir_od,"Data/era5")
dir_haduk<-file.path(dir_od,"Data/HadUK/2024")

# File pathways
dtm_file<-file.path(dir_dtm,"uk_dtm.tif")
tree_file<-file.path(dir_field,"Field_data_locations.csv")
site_file<-file.path(dir_field,"site_properties.csv") # general properties for modelling nsp at sites
grid_file<-file.path(dir_od,'Data','OSgrids','os_bng_grids.gpkg')

file.exists(dtm_file);file.exists(tree_file);file.exists(site_file); file.exists(grid_file)
```


# Location of field sites

```{r data_prep, include=FALSE}
# Location file (by tree)
tree_locs<-read.csv(tree_file)
site_props<-read.csv(site_file,strip.white = TRUE )
sites<-unique(site_props$site,)

# Site & tree geo data
trees_sf<-st_as_sf(tree_locs,coords = c("lon", "lat"), crs = 4326)
sites_sf<-st_as_sf(site_props,coords = c("lon", "lat"), crs = 4326)

# Get site terrain data and extract slope, elev and aspect
sites_27700<-st_transform(sites_sf,27700)
e<-ext(st_bbox(sites_27700)) # extent of area with sites
dtm<-rast(dtm_file)

#terrain_r<-terrain(dtm,v=c("slope","aspect"))
terrain_r<-rast(file.path(dir_dtm,"uk_terrain.tif"))
elev<-terra::extract(dtm,sites_27700)
slope<-terra::extract(terrain_r[["slope"]],sites_27700)
aspect<-terra::extract(terrain_r[["aspect"]],sites_27700)
sites_sf<-cbind(sites_sf,elev=elev$NH35,slope=slope$slope,aspect=aspect$aspect)

# If required calculate site lat lon from mean of the trees there
# sites<-data.frame(site=unique(locs$Site[order(locs$Site)]),lon=tapply(locs[,c("lon")],as.factor(locs$Site),mean),lat=tapply(locs[,c("lat")],as.factor(locs$Site),mean))
```


```{r sitemap}
centre_pt<-as.numeric(st_coordinates(st_centroid(st_as_sfc(st_bbox(sites_sf)))))
popup_text <- paste("<strong>Tree ID: </strong>",  trees_sf$Site, trees_sf$Tree_species, trees_sf$Tree)
leaflet(options=list(minZoom=4)) %>%
  setView(lng = centre_pt[1], lat = centre_pt[2], zoom = 7) %>%
  addProviderTiles(providers$OpenStreetMap.Mapnik,group="Streetmap") %>%
  addCircleMarkers(data=trees_sf, 
             popup = popup_text, 
             color = ~ ifelse(Tree_species == "oak", "red", "blue"))
```

# Load climate temperature data for sites
See Get_2024_weather.Rmd for preprocessing of 2024 Haduk data.

During the main monitoring period (Feb to Oct), temperature were not lower than -3C and no higher than 33C.

```{r,2024_weather}
temp<-rast(file.path(dir_od,"MetOff_data","weather2024","temp_hrly_2024.tif"))
plot(temp[[1]]); plot(sites_27700,add=T,col='red')
tme<-time(temp)
t2m<-t(extract(temp,vect(sites_27700),method="simple",ID=TRUE))

t2m_df<-as.data.frame(t2m[2:dim(t2m)[1],],rownames=NULL)
names(t2m_df)<-tolower(sites_sf$site)
t2m_df$tme<-time(temp)
rownames(t2m_df)<-NULL

# Plot range of temperature values in monitoring period (Feb-Oct)
sel<-which(t2m_df$tme>=as.POSIXct("2024-02-01") & t2m_df$tme<=as.POSIXct("2024-10-10"))
for(ste in c('ah','be','cw','kg','wh')){
  print(paste(ste,round(min(t2m_df$ah[sel]),2),'to',round(max(t2m_df$ah[sel]),2),'C'))
}

```

# Load logger temperature readings 
Site start/end dates:

AH    NSP   17/02-19/10
      OAK   17/02-22/09
      
BE    ALL   14/02-01/10

CW    NSP   18/03-24/09
      OAK1&2 03/04-11/10
      OAK3  27/04-11/10
      
KG    ALL   01/03-18/09

WH    ALL   22/03-02/10


## Load surveytag downloads and organise

```{r load_surveytags}
#Survey tags
stfiles<-list.files(dir_surveytag, pattern =".csv",recursive=TRUE,full.names=TRUE)

# Extract metadata from file names and create id variable for matching files by tag/location
for(f in stfiles){
    metadata<-as.data.frame(get_surveytag_metadata(f))
    metadata<-cbind(filepath=f,metadata)
    if (f==stfiles[1]) file_metadata<-metadata else file_metadata<-rbind(file_metadata,metadata)
}

file_metadata$id=tolower(paste(file_metadata$site,file_metadata$spp,file_metadata$tree,sep='_'))
head(file_metadata); nrow(file_metadata)
View(file_metadata)

# Number of files per id (tree) 
table(file_metadata$id) # =5 to 9 files per tree

# Number of different loggers per tree? BUt some loggers appear named the same?
loggernum_df<-file_metadata %>%
  group_by(id) %>%
  summarize(logger_num = n_distinct(logger))
View(loggernum_df)


### Load and merge all files by tree id while performing quality control checks
ids<-unique(file_metadata$id)

columns <- c("id","site","spp","tree","type","datetime", "air_temp") 
stagdata_df<- data.frame(matrix(nrow = 0, ncol = length(columns))) 
colnames(stagdata_df) <- columns

for(id in ids){ #  For each tree ID...
  alldat<- data.frame(matrix(nrow = 0, ncol = length(columns))) 
  colnames(alldat) = columns

  #print(paste("Processing files for",id))
  iddat_df<-file_metadata[which(file_metadata$id==id),]
  # Check if info in file
  finfo<-file.info(iddat_df$filepath)
  ignore<-which(is.na(finfo$size))
  if(length(ignore)>0){
    warning(paste("Ignoring file",iddat_df$filepath[ignore]))
    iddat_df<- iddat_df[-ignore,]
  }

  for(f in iddat_df$filepath){
    tagdat<-load_surveytag(f,rnd30=FALSE)
    tagdat$id<-id
    alldat<-rbind(alldat,tagdat)
  }
  alldat<-alldat[order(alldat$datetime),]
  
  # Remove readings of 0.00000 as should be NA
  alldat$air_temp<-ifelse(alldat$air_temp==0.00000,NA,alldat$air_temp)
  
  # Remove duplicate timed readings
  dup<-duplicated(alldat$datetime)
  alldat<-alldat[!dup,]
  
  # Remove excessively high/low readings 
  minlimit<- -5
  maxlimit<- 40
  extrmT<-which(alldat$air_temp < minlimit | alldat$air_temp > maxlimit )
  if(length(extrmT)>0) {
    warning(paste('Removing extreme reading of',alldat$air_temp[extrmT], 'recorded by logger',alldat$id[extrmT],',   ') )
    alldat<-alldat[-extrmT,]
  }
  
  # Remove very isolated readings (no other within 1 hour) 
  isolated_idx<-find_isolated_times(alldat$datetime,limit=3600)
  if(length(isolated_idx>0)){ 
    warning(paste('Removing',length(isolated_idx), 'isolated readings from' ,id))
    alldat<-alldat[-isolated_idx,]
  }
  # Join all location data in single df
  stagdata_df<-rbind(stagdata_df,alldat)
}

# Save results
saveRDS(stagdata_df,file.path(dir_out,'surveytag_air_logger_results.Rds'))
```

### Plot surveytags by site and tree
General issues for sites:

AH - 
BE - poor and erratic reliability during start of season
CW - late start, NSP mid march, Oaks April
KG - ~one month missing data for ALL trees 9th to 29th May ??
WH - ~two months missing for ALL trees 24 May to 21 June & 24 Sept to 21 August??

Loggers show several areas that look like spurious readings particularly at start or end of time series.
Other larger periods of times where logge strays from other readings at site include be-nsp-5

```{r ggplot_surveytags}
stagdata_df<-readRDS(file.path(dir_out,'surveytag_air_logger_results.Rds'))

# Plot temp timeseries
minT<-min(stagdata_df$air_temp,na.rm=T)
maxT<-max(stagdata_df$air_temp,na.rm=T)
mintme<-min(stagdata_df$datetime)
maxtme<-max(stagdata_df$datetime)


for(ste in unique(stagdata_df$site)){
  plotdata<-stagdata_df[which(stagdata_df$site==ste),]
  allids <- unique(plotdata$id)
  allids<-substr(allids,4,8)
  oak_cols<-c('darkred','red','orange','pink','brown','yellow')
  nsp_cols<-c('darkblue','blue','lightblue','cyan','purple','skyblue')
  
  no<-1;
  nn<-1;
  gcolors<-c()
  for(i in allids){
    if(substr(i,1,3)=='oak'){
      gcolors<-c(gcolors,oak_cols[no])
      no<-no+1
      }
    if(substr(i,1,3)=='nsp'){ 
      gcolors<-c(gcolors,nsp_cols[nn])
      nn<-nn+1
    }
  }

  fig<-ggplot(data = plotdata, aes(x = datetime, y = air_temp, colour = id)) + 
    ggtitle(paste("Site ",ste)) + 
    ylim(minT,maxT) +
    xlab("Date") + 
    ylab(paste('Air temperature')) + 
    geom_line(aes(group = id))+
    scale_colour_manual(values=gcolors)+
    scale_x_datetime(date_breaks = "months" , date_labels = "%b", limits=c(mintme,maxtme))
  print(fig)
}
```

## Load Tinytag data
Input from output of Tinytag software Save Table as UTF no units .csv file.
Assume black probe = deeper/inner; white probe = shallow/outer locations in trunk.

Individual observations:
AH oak-1-s  black probe sudden large increase (~5C) from 17/07 - REMOVE!!!
AH oak-4-s black probe slight increase 30/06
AH oak-5-s  black probe slight increase 02/09

BE nsp-6-n Black probe large decline between 26*06-18/07 - REMOVE!!!

CW oak-2-s black probe sudden large increase (~5C) from 15/07 - REMOVE!!!
CW oak-2-n black probe slight increase from 22/09
CW oak-3-n&s Starts 26/04

WH nsp-2-s Starts 29/05
WH nsp-4-n & s Black probe -40 until 18/04 13h then normal - REMOVE!!!
WH oaks - generally white probe shows slight increase during summer then maintained (eg ~1-2C > black)
WH-oak-3-s spike black spike 09/08 13h - 10/08 02h - REMOVE!!

```{r load_multiple_tinytags}
ttfiles<-list.files(dir_tinytag,pattern =".csv",recursive=TRUE,full.names=TRUE)
#ttfiles<-ttfiles[which(substr(basename(ttfiles),1,2) %in% site_id)]

# Results for all files and ids
columns <- c("id","site","spp","tree","asp","type","datetime", "inner_temp","outer_temp") 
ttagdata_df<- data.frame(matrix(nrow = 0, ncol = length(columns))) 
colnames(ttagdata_df) = columns


# Extract metadata from file names and create id variable for matching files by tag/location
for(f in ttfiles){
    metadata<-as.data.frame(get_tinytag_metadata(f))
    metadata<-cbind(filepath=f,metadata)
    if (f==ttfiles[1]) file_metadata<-metadata else file_metadata<-rbind(file_metadata,metadata)
}

file_metadata$id=tolower(paste(file_metadata$site,file_metadata$spp,file_metadata$tree,sep='_'))

## Can now load and merge files by id
ids<-unique(file_metadata$id)

for(id in ids){
  # id results df
  alldat<- data.frame(matrix(nrow = 0, ncol = length(columns))) 
  colnames(alldat) = columns

  iddat_df<-file_metadata[which(file_metadata$id==id),]
  for(f in iddat_df$filepath){
    tagdat<-load_tinytag(f,rnd30=FALSE)
    tagdat$id<-id
    if(f==iddat_df$filepath[1]) alldat<-tagdat else alldat<-rbind(alldat,tagdat)
  }
  alldat<-alldat[order(alldat$datetime),]
  
  # Remove duplicate readings (but only if temp and datetime identical!)
  dup<-duplicated(alldat)
  alldat<-alldat[!dup,]
  
  # Join all location data in single df
  ttagdata_df<-rbind(ttagdata_df,alldat)
}

# Remove excessively high/low readings 
minlimit<- -10
maxlimit<- 45
extrmT<-which(ttagdata_df$inner_temp < minlimit | ttagdata_df$inner_temp > maxlimit )
if(length(extrmT)>0) {
  warning(paste('Removing extreme reading of',ttagdata_df$inner_temp[extrmT], 'recorded by logger',ttagdata_df$id[extrmT],',   ') )
  ttagdata_df$inner_temp[extrmT]<-NA
}
extrmT<-which(ttagdata_df$outer_temp < minlimit | ttagdata_df$outer_temp > maxlimit )
if(length(extrmT)>0) {
  warning(paste('Removing extreme reading of',ttagdata_df$outer_temp[extrmT], 'recorded by logger',ttagdata_df$id[extrmT],',   ') )
  ttagdata_df$outer_temp[extrmT]<-NA
}


# MANUAL CORRECTION FROM OBSERVATIONS to Black probe readings
# ah-oak-1-s Black probe 17/07 6h to end
sel<-which(ttagdata_df$fullid=='ah_oak_1_s' & ttagdata_df$datetime>=as.POSIXct("2024-07-17 06:00"))
ttagdata_df$inner_temp[sel]<-NA

# be-nsp-6-n Black probe 27/06 17h to 17/07 19.30
sel<-which(ttagdata_df$fullid=='be_nsp_6_n' & ttagdata_df$datetime>as.POSIXct("2024-06-27 11:00") & ttagdata_df$datetime<as.POSIXct("2024-07-17 19:30"))
ttagdata_df$inner_temp[sel]<-NA

# cw-oak-2-s Black probe 15/07 7h  to end
sel<-which(ttagdata_df$fullid=='cw_oak_2_s' & ttagdata_df$datetime>=as.POSIXct("2024-07-15 07:00"))
ttagdata_df$inner_temp[sel]<-NA

# wh-oak-3-s Black probe 08/08 22h to 10/08 5h
sel<-which(ttagdata_df$fullid=='wh_oak_3_s' & ttagdata_df$datetime>as.POSIXct("2024-08-08 22:00") & ttagdata_df$datetime<as.POSIXct("2024-08-10 05:00"))
ttagdata_df$inner_temp[sel]<-NA

# Save results
saveRDS(ttagdata_df,file.path(dir_out,'tinytag_bark_logger_results.Rds'))

```

### GGplot of ttag data by tree
```{r ggplot_ttag_bytree, eval=FALSE}
# Plot temp timeseries
minT<-min(ttagdata_df[,c("inner_temp","outer_temp")],na.rm=T)
maxT<-max(ttagdata_df[,c("inner_temp","outer_temp")],na.rm=T)
mintme<-min(ttagdata_df$datetime)
maxtme<-max(ttagdata_df$datetime)

gcolors <- c("s_inner" = "red", "s_outer" = "orange", "n_inner"="blue", "n_outer" = "cyan")
save<-FALSE

for(id in unique(ttagdata_df$id)){
  data<-ttagdata_df[which(ttagdata_df$id==id),]
  inner<-data[,c('id','site','spp','tree','asp','type','datetime','inner_temp')]
  outer<-data[,c('id','site','spp','tree','asp','type','datetime','outer_temp')]
  names(inner)<-c(names(inner)[1:7],'temperature')
  names(outer)<-c(names(outer)[1:7],'temperature')
  inner$depth<-'inner'
  outer$depth<-'outer'
  plotdata<-rbind(inner,outer)
  plotdata$location<-paste0(plotdata$asp,'_',plotdata$depth)
  
  
  fig<-ggplot(data = plotdata, aes(x = datetime, y = temperature, colour = location)) + 
    ggtitle(paste("Tree ",plotdata$id[1])) + 
    ylim(minT,maxT) +
    xlab("Date") + 
    ylab(paste('Temperature')) + 
    geom_line(aes(group = location))+
    scale_colour_manual(values=gcolors)+
    scale_x_datetime(date_breaks = "months" , date_labels = "%b", limits=c(mintme,maxtme))
  
  if(save){
    figname<-paste0(id,"_trunk_temp_timeseries.png")
    ggsave(file.path(dir_out,figname),fig)
  }
  print(fig)
}
```

# Merge weather, air and trunk temperatures
Merges timeseries - creates regular timeseries of readings for weather, air and bark temperatures.

* Linear interpolates temperatures to regular time intervalsof 30mins.

* Does NOT interpolate if no reading within 30mins.

```{r merge_temps}
# Order observations by datetime 
stagdata_df<-stagdata_df[order(stagdata_df$id, stagdata_df$datetime),]
ttagdata_df<-ttagdata_df[order(ttagdata_df$id,ttagdata_df$datetime),]

# Create regular datetime series
dt<-c(stagdata_df$datetime,ttagdata_df$datetime)
dt_series<-regular_timeseries(dt,interval="30 mins")

# Define results df
colnames<-c("id", "site","spp","tree","datetime","air_temp","n_outer_temp","n_inner_temp","s_outer_temp","s_inner_temp")
treeresults_df<-data.frame(matrix(NA,nrow=0,ncol=length(colnames)))
names(treeresults_df)<-colnames


# By tree id, merge and interpolate logger readings to common, regular time series
# Define max time step for interpolating
tme_thold<-1+(60*60) # 30mins in seconds

for(tree_id in unique(stagdata_df$id)){
  #print(tree_id)
  # Get data for tree
  sel<-which(stagdata_df$id==tree_id)
  svydat<-stagdata_df[sel,]
  ste<-unique(svydat$site)
  
  # Two sets of tinytag data - norht and south
  sel<-which(ttagdata_df$id==tree_id & ttagdata_df$asp=="s")
  ttgdat_s<-ttagdata_df[sel,]
  sel<-which(ttagdata_df$id==tree_id & ttagdata_df$asp=="n")
  ttgdat_n<-ttagdata_df[sel,]

  results_df<-data.frame(
    id=rep(unique(svydat$id),length(dt_series)),
    site=rep(unique(svydat$site),length(dt_series)),
    spp=rep(unique(svydat$spp),length(dt_series)),
    tree=rep(unique(svydat$tree),length(dt_series)),
    datetime=dt_series,
    weather_temp=match_temperatures(t2m_df[,ste], t2m_df$tme, dt_series,tme_thold),
    air_temp=match_temperatures(svydat$air_temp, svydat$datetime, dt_series,tme_thold),
    n_outer_temp=match_temperatures(ttgdat_n$outer_temp, ttgdat_n$datetime, dt_series,tme_thold),
    n_inner_temp=match_temperatures(ttgdat_n$inner_temp, ttgdat_n$datetime, dt_series,tme_thold),
    s_outer_temp=match_temperatures(ttgdat_s$outer_temp, ttgdat_s$datetime, dt_series,tme_thold),
    s_inner_temp=match_temperatures(ttgdat_s$inner_temp, ttgdat_s$datetime, dt_series,tme_thold)
   )
  # Collate into single df
  if(tree_id == unique(stagdata_df$id)[1]) allresults_df<-results_df else allresults_df<-rbind(allresults_df,results_df)
}

saveRDS(allresults_df,file.path(dir_out,'logger_results.Rds'))
```


## Merge with modelled LAI data from hemisfer measurements and photos
See:
Values from fitting models to LAI observations derived from hemispherical photographs.
Only available for AH, BE, CW & WH sites.


```{r merge_lai}
lai_type<-'lc'
# Fitted models file
lai_models<-readRDS(file.path(dir_lai,paste0(lai_type,'_lai_models.Rds')))

allresults_df<-readRDS(file.path(dir_out,'logger_results.Rds'))

allresults_df$lai<-NA
for(t in names(lai_models)){
  model<-lai_models[[t]] 
  sel<-which(allresults_df$id==t)
  # Get day of year from result datetime
  doy<-lubridate::yday(allresults_df$datetime[sel])
  pred_lai<-predict(model, data.frame(date=doy))
  #print(plot(pred_lai~doy,main=t))
  allresults_df$lai[sel]<-round(pred_lai,2)
  #print(plot(allresults_df$lai[sel],main=t))
}
```

## Compare air and weather temperatures

### By site and manual editing of ait temperature data
Dynamic graph for each site showing each tree air temperature and site weather temperature.
Option: calculate average site air temp time series to infill where missing values? But KG and WH missing all logger readings for certain periods of time.

Notes:
be_nsp_5 survey tag recoridng about 5C lower than all others
be_oak_1  survey tag show erratic behaviour - strings of 0s and NAs
be_oak_3 - erratic - some 0.000 readings just before string of NAs


```{r plot_air_weather_bysite}
allresults_df<-readRDS(file.path(dir_out,'logger_results.Rds'))
for(ste in unique(allresults_df$site)){
  sitedata<-allresults_df[which(allresults_df$site ==ste),c("id","site","spp","tree","datetime","weather_temp", "air_temp")]
  airdata<-as.data.frame(sitedata %>% 
                           pivot_wider(id_cols=datetime,names_from=id,names_glue = "{id}_air", values_from=air_temp))
  weatherdata<-sitedata$weather_temp[which(sitedata$id==unique(sitedata$id)[1])]
  length(weatherdata)==nrow(airdata)
  plotdata<-cbind(airdata,weather=weatherdata)
  tseries<- xts(x = plotdata[,-1], order.by = plotdata$datetime)
  ts_names<-names(tseries[,-1])
  oak_names<-ts_names[which(substr(ts_names,4,6)=='oak')]
  nsp_names<-ts_names[which(substr(ts_names,4,6)=='nsp')]
  
  dynfig <- dygraph(tseries) %>% dyAxis("y", label = "Temp (C)")
  dynfig<-dynfig %>% dySeries("weather",color = 'black',strokeWidth = 1.5) 
  for (nme in oak_names){
    dynfig<-dynfig %>% dySeries(nme,color='purple')
  }
  for (nme in nsp_names){
    dynfig<-dynfig %>% dySeries(nme,color='green')
  }
  print(dynfig)
    #dyHighlight(highlightSeriesOpts = list(strokeWidth = 3))
  htmltools::tagList(dynfig)
}
```
Manual edits described in file applied and saved.

```{r manual_edits}
allresults_df<-readRDS(file.path(dir_out,'logger_results.Rds'))
edits_file<-file.path(dir_field,"Logger_results","manual_airtemp_edits.csv")
edits_df<-read.csv(edits_file)[,1:3]

edits_df$start<-as.POSIXct(trim(edits_df$start),"%d/%m/%Y %H:%M", tz=tz(allresults_df$datetime))
sel<-which(edits_df$start>as.POSIXct("31/03/2024 01:00","%d/%m/%Y %H:%M", tz=tz(allresults_df$datetime)))
edits_df$start[sel]<-edits_df$start[sel]-60*60

edits_df$end<-as.POSIXct(edits_df$end,"%d/%m/%Y %H:%M", tz=tz(allresults_df$datetime))
sel<-which(edits_df$end>as.POSIXct("31/03/2024 01:00","%d/%m/%Y %H:%M", tz=tz(allresults_df$datetime)))
edits_df$end[sel]<-edits_df$end[sel]-60*60


edits_df
for(n in 1:nrow(edits_df)){
  sel<-which(
    allresults_df$id==edits_df$id[n] &
    as.numeric(allresults_df$datetime) >= as.numeric(edits_df$start[n]) &
    as.numeric(allresults_df$datetime) <= as.numeric(edits_df$end[n]) )
  allresults_df$air_temp[sel]<-NA
}

saveRDS(allresults_df,file.path(dir_out,'logger_results_stedit.Rds'))
```

Ah - summer buffering
BE - more extreme day and night than weather
CW - some daytime buffer in summer
KG - low variation in air temps and from weather temp
WH - some daytime buffering in summer

### By tree
Plot air and weather by individual tree - helps identify spurious survey tag readings

```{r plot_air_weather_bytree, eval=FALSE}
allresults_df<-readRDS(file.path(dir_out,'logger_results_stedit.Rds'))
for(id in unique(allresults_df$id)){
  plotdata<-allresults_df[which(allresults_df$id ==id),c("id","site","spp","tree","datetime","weather_temp", "air_temp")]
  tseries<- xts(x = plotdata[,c('weather_temp','air_temp')], order.by = plotdata$datetime)
  dynfig <- dygraph(tseries) %>% dyAxis("y", label = "Temp (C)")
  print(dynfig)
}
```

## Summarise air/weather differences by site and species
Calculate temperature difference as: air temperature - weather temperature for each tree (-ve = where air temperature lower than weather).
Plots distributions of difference for each site/species combination (ignoring values + or -10C or more).
```{r hist_air_weather_difs}
difdata<-allresults_df[which(!is.na(allresults_df$weather_temp) &  !is.na(allresults_df$air_temp)), c("id","site","spp","tree","datetime","weather_temp", "air_temp")]

# +ve when air temp>climate (night) -ve if air temp < weather (day)
difdata$diff_temp<-difdata$air_temp-difdata$weather_temp

# Remove extrem values +/-10 or more
difdata<-difdata[-which(difdata$diff_temp>10 | difdata$diff_temp< -10),]

# Noxplot / histogram of difference freq by site/spp (season?) - 5x2 GRID plot
par(mfrow=c(5,2))
par(mar=c(1,1,1,1))
# Calculate limits
for (ste in unique(difdata$site)){
  for (spp in unique(difdata$spp)){
    sel<-which(difdata$site==ste & difdata$spp==spp)
    plotdata<-difdata[sel,]
    hist(plotdata$diff_temp,main=paste(ste,spp),xlim=c(-10,10), breaks=seq(-10,10,1))
  }
}

```


## Plot all data by tree: air, bark and weather temperatures
In general inner and outer show buffered variation compared with air temp EXCEPT for certain times - where much higher peaks - due to direct radiation on trunks?
Very slight phase shift when trunk temps bufferd ()

Alice Holt
Follow general pattern - buffered inner bark temperatures.
ah_oak_3 example of elevated outer bark temperatures (sometimes greater than weather/canopy air temperatures) - cf with ah_oak_4 much more buffered bark temps

Example wh_oak_5 - matches expected pattener of variation across weather, air, bark 


NSP6 : very different n-facing trunk temperatures


```{r plot_timeseries_bytree}
# Plot temp timeseries
allresults_df<-readRDS(file.path(dir_out,'logger_results_stedit.Rds'))

for(id in unique(allresults_df$id)){
  plotdata<-allresults_df[which(allresults_df$id ==id),
                c("datetime","weather_temp", "air_temp",
                  "n_outer_temp", "n_inner_temp", "s_outer_temp", "s_inner_temp")]
  
  tseries<- xts(x = plotdata, order.by = plotdata$datetime)

  dynfig <- dygraph(tseries, main=id) %>% dyAxis("y", label = "Temp (C)")
  dynfig<-dynfig %>% dySeries("weather_temp",color = 'black') 
  dynfig<-dynfig %>% dySeries("air_temp",color = 'green') 
  dynfig<-dynfig %>% dySeries("n_outer_temp",color = 'cyan') 
  dynfig<-dynfig %>% dySeries("n_inner_temp",color = 'blue') 
  dynfig<-dynfig %>% dySeries("s_outer_temp",color = 'orange') 
  dynfig<-dynfig %>% dySeries("s_inner_temp",color = 'red') 
  dynfig<-dynfig %>% dyLegend(width=600)
  #dynfig<-dynfig %>% dyLegend(show="follow",labelsSeparateLines=TRUE)
  
  show(dynfig)

}
```


# CUT OUTS

### Plots of Inner and outer temperatures for each tree and location on trunk
```{r plot_bark locations, eval=FALSE}
ttagdata_df<-readRDS(file.path(dir_out,'tinytag_bark_logger_results.Rds'))

# Plot temp timeseries by tree and aspect
minT<-min(ttagdata_df[,c("inner_temp","outer_temp")],na.rm=T)
maxT<-max(ttagdata_df[,c("inner_temp","outer_temp")],na.rm=T)

for(fullid in unique(ttagdata_df$fullid)){
  print(fullid)
  plotdata<-ttagdata_df[which(ttagdata_df$fullid==fullid),]
  plot_tseries(plotdata,title=fullid,ylimits=c(minT,maxT))
}
```

### Plot differences in bark temperatures
Calculate difference aas outer - inner bark temperatures for 
A few anomalous differences
ah-nsp-1-n - two brief dips to ~-5
```{r plot_differences, eval=FALSE}
ttagdata_df<-readRDS(file.path(dir_out,'tinytag_bark_logger_results.Rds'))

ttagdata_df$dif_temp<-ttagdata_df$outer_temp-ttagdata_df$inner_temp
minT<-min(ttagdata_df$dif_temp,na.rm=T)
maxT<-max(ttagdata_df$dif_temp,na.rm=T)
mintme<-min(ttagdata_df$datetime,na.rm=T)
maxtme<-max(ttagdata_df$datetime,na.rm=T)

for(fullid in unique(ttagdata_df$fullid)){
  plotdata<-ttagdata_df[which(ttagdata_df$fullid==fullid),c('fullid','datetime','dif_temp')]
  plot_tseries(plotdata,title=paste(fullid,": Outer - Inner bark difference (C)"), ylimits=c(minT,maxT),xlimits=c(mintme,maxtme))
  abline(h=0)
}
```
