---
title: "Data Processing"
---
Refs
MyClim: 
https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.14192
http://labgis.ibot.cas.cz/myclim/index.html
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(devtools)
library(stringr)
library(terra)
library(ncdf4)
dir_root<-"/Users/jonathanmosedale/Library/CloudStorage/OneDrive-UniversityofExeter/Add_Trees"
site_id<-"cw"
dir_tinytag<-dir_root
dir_surveytag<-dir_root

# Restrict to single site eg be, cw
site_id<-"cw"
```

## Load Surveytag data
Recorded air temperature as TC_Reading in .csv output of surveytags.

```{r load_surveytag}
stfiles<-list.files(dir_surveytag,pattern =".csv",recursive=TRUE,full.names=TRUE)
stfiles<-stfiles[which(substr(basename(stfiles),1,2) %in% site_id)]
f<-stfiles[38]
stdata_df<-load_surveytag(f,rnd30=FALSE)
plot_tseries(stdata_df)

f<-stfiles[37]
stdata_df<-load_surveytag(f,rnd30=FALSE)
plot_tseries(stdata_df)

```


## Load multiple surveytag downloads and organise
cw_nsp4 has llots of duplicate readings...
```{r load_multitags}
#Survey tags
stfiles<-list.files(dir_surveytag,pattern =".csv",recursive=TRUE,full.names=TRUE)
stfiles<-stfiles[which(substr(basename(stfiles),1,2) %in% site_id)]

print(stfiles)

# Extract metadata from file names and create id variable for matching files by tag/location
for(f in stfiles){
    metadata<-as.data.frame(get_surveytag_metadata(f))
    metadata<-cbind(filepath=f,metadata)
    if (f==stfiles[1]) file_metadata<-metadata else file_metadata<-rbind(file_metadata,metadata)
}

file_metadata$id=tolower(paste(file_metadata$site,file_metadata$spp,file_metadata$tree,sep='_'))
head(file_metadata); nrow(file_metadata)

## Can now load and merge files by id

ids<-unique(file_metadata$id)
ids<-ids[which(substr(ids,1,2) %in% site_id)]
for(id in ids){ #  For each logger...
  iddat_df<-file_metadata[which(file_metadata$id==id),]
  for(f in iddat_df$filepath){
    tagdat<-load_surveytag(f,rnd30=FALSE)
    tagdat$id<-id
    if(f==iddat_df$filepath[1]) alldat<-tagdat else alldat<-rbind(alldat,tagdat)
  }
  alldat<-alldat[order(alldat$datetime),]
  
  # Remove duplicate readings
  dup<-duplicated(alldat)
  alldat<-alldat[!dup,]
  
  # Remove readings of 0.00000 as should be NA
  alldat$air_temp<-ifelse(alldat$air_temp==0.00000,NA,alldat$air_temp)
  
  # Remove duplicate datetimes
  dp<-which(duplicated(alldat$datetime))
  if(length(dp)>0) {
    print(paste('Duplicate readings found in file...',f))
    for(d in dp){ # for each duplicate
      cpy<-which(alldat$datetime==alldat$datetime[d]) # get all copies
      avgvals<-alldat$air_temp[c(min(cpy)-1,max(cpy+1))] # calculate avg temp before and after
      keep<-which(alldat$air_temp[cpy]-avgvals==min(alldat$air_temp[cpy]-avgvals)) # keep value closest to above
      remove_row<-cpy[-keep]
      alldat<-alldat[-remove_row,]
    }
  }
  
  # Join all location data in single df
  if(id==ids[1]) stagdata_df<-alldat else stagdata_df<-rbind(stagdata_df,alldat)
}

for(id in unique(stagdata_df$id)){
  print(id)
  plotdata<-stagdata_df[which(stagdata_df$id==id),]
  plot_tseries(plotdata,title=id)
}
```
## Get era5 data
```{r era5}
era5file<-file.path(dir_root,"era5_2024.nc")
t2m<-rast(era5file,"t2m")
tmp<-extract(t2m,c(50.33,-4.98))




## Load Tinytag data
Input must be from Tinytag software Save Table as UTF no units .csv file.
Assume black probe = deeper/inner and white probe = shallow/outer locations in trunk.

```{r load_tinytag}
ttfiles<-list.files(dir_tinytag,pattern =".csv",recursive=TRUE,full.names=TRUE)
ttfiles<-ttfiles[which(substr(basename(ttfiles),1,2) %in% site_id)]

#print(ttfiles)
f<-ttfiles[1]
ttdata_df<-load_tinytag(f)
plot_tseries(ttdata_df)
```


# Load multiple tinytag csv files
```{r load_multiple_tinytags}
ttfiles<-list.files(dir_tinytag,pattern =".csv",recursive=TRUE,full.names=TRUE)
ttfiles<-ttfiles[which(substr(basename(ttfiles),1,2) %in% site_id)]

# Extract metadata from file names and create id variable for matching files by tag/location
for(f in ttfiles){
    metadata<-as.data.frame(get_tinytag_metadata(f))
    metadata<-cbind(filepath=f,metadata)
    if (f==ttfiles[1]) file_metadata<-metadata else file_metadata<-rbind(file_metadata,metadata)
}

file_metadata$id=tolower(paste(file_metadata$site,file_metadata$spp,file_metadata$tree,sep='_'))

## Can now load and merge files by id
ids<-unique(file_metadata$id)

for(id in ids){
  iddat_df<-file_metadata[which(file_metadata$id==id),]
  for(f in iddat_df$filepath){
    tagdat<-load_tinytag(f,rnd30=FALSE)
    tagdat$id<-id
    if(f==iddat_df$filepath[1]) alldat<-tagdat else alldat<-rbind(alldat,tagdat)
  }
  alldat<-alldat[order(alldat$datetime),]
  
  # Remove duplicate readings (but only if temp and datetime identical!)
  dup<-duplicated(alldat)
  alldat<-alldat[!dup,]
  # Join all location data in single df
  if(id==ids[1]) ttagdata_df<-alldat else ttagdata_df<-rbind(ttagdata_df,alldat)
}

# Plot temp timeseries
for(fullid in unique(ttagdata_df$fullid)){
  print(fullid)
  plotdata<-ttagdata_df[which(ttagdata_df$fullid==fullid),]
  plot_tseries(plotdata,title=fullid)
}
```



## Merge air and trunk temperatures
Question: does it require merging timeseries - ie selecting closest air temp for each trunk temp reading?
Interpolate air temp to trunk temp times

Should NOT interpolate if no air temp within 60mins

If air measurements within ? time then interpolate
If no air measurements within then keep blank

```{r merge_temps}
# Order observations by datetime 
stagdata_df<-stagdata_df[order(stagdata_df$id, stagdata_df$datetime),]
ttagdata_df<-ttagdata_df[order(ttagdata_df$id,ttagdata_df$datetime),]

# Create regular datetime series
dt<-c(stagdata_df$datetime,ttagdata_df$datetime)
dt_series<-regular_timeseries(dt,interval="30 mins")

# Define results df
colnames<-c("id", "site","spp","tree","datetime","air_temp","n_outer_temp","n_inner_temp","s_outer_temp","s_inner_temp")
treeresults_df<-data.frame(matrix(NA,nrow=0,ncol=length(colnames)))
names(treeresults_df)<-colnames


# By tree id, merge and interpolate logger readings to common, regular time series
# Define max time step for interpolating
tme_thold<-1+(60*60) # 30mins in seconds

for(tree_id in unique(stagdata_df$id)){
  print(tree_id)
  # Get data for tree
  sel<-which(stagdata_df$id==tree_id)
  svydat<-stagdata_df[sel,]
  # Two sets of tinytag data - norht and south
  sel<-which(ttagdata_df$id==tree_id & ttagdata_df$asp=="s")
  ttgdat_s<-ttagdata_df[sel,]
  sel<-which(ttagdata_df$id==tree_id & ttagdata_df$asp=="n")
  ttgdat_n<-ttagdata_df[sel,]

  results_df<-data.frame(
    id=rep(unique(svydat$id),length(dt_series)),
    site=rep(unique(svydat$site),length(dt_series)),
    spp=rep(unique(svydat$spp),length(dt_series)),
    tree=rep(unique(svydat$tree),length(dt_series)),
    datetime=dt_series,
    air_temp=match_temperatures(svydat$air_temp, svydat$datetime, dt_series,tme_thold),
    n_outer_temp=match_temperatures(ttgdat_n$outer_temp, ttgdat_n$datetime, dt_series,tme_thold),
    n_inner_temp=match_temperatures(ttgdat_n$inner_temp, ttgdat_n$datetime, dt_series,tme_thold),
    s_outer_temp=match_temperatures(ttgdat_s$outer_temp, ttgdat_s$datetime, dt_series,tme_thold),
    s_inner_temp=match_temperatures(ttgdat_s$inner_temp, ttgdat_s$datetime, dt_series,tme_thold)
   )
  # Collate into single df
  if(tree_id == unique(stagdata_df$id)[1]) allresults_df<-results_df else allresults_df<-rbind(allresults_df,results_df)
}

```


## Create site timeseries of average air_temperatures
Might be used to infill where loggers not functioning for long period of time

```{r svytag_avg_timeseries}
site_id<-"be"
airdat<-allresults_df[which(allresults_df$site==site_id),c("id","site","spp","tree","datetime","air_temp")]
tree_airtemp<-as.data.frame(airdat %>% pivot_wider(id_cols=datetime,names_from=id,names_glue = "{id}_temp", values_from=air_temp))
names(tree_airtemp)

plot_tseries(tree_airtemp,title='Air temp',legloc='topleft',cols=c('cyan','blue','purple','violet','green','skyblue', 'red','orange','yellow'))


ggplot(allresults_df,aes(x=datetime,y=air_temp)) +
  stat_summary(fun.data ="mean_sdl", mult=1, geom = "smooth")

## OR use gg[plot!!!
library(ggplot2)
ggplot(data=allresults_df, aes(x=datetime, y=air_temp, group = id, colour = id)) + 
    geom_line() +labs(y= "Air temperature loggers", x = "Date") +
    theme(legend.key.size = unit(2,"line"))
  #stat_summary(fun.y = mean,  colour = "darkgrey")
  stat_summary(fun.data ="mean_sdl", mult=1, geom = "smooth")
#


  # Short time series
startdt<-as.POSIXct(strptime("2024-03-10 00:00:00","%Y-%m-%d %H:%M:%S"), tz='GMT')
enddt<-as.POSIXct(strptime("2024-04-10 00:00:00","%Y-%m-%d %H:%M:%S"), tz='GMT')
ggplot(data=allresults_df[which(allresults_df$datetime>=startdt & allresults_df$datetime<=enddt),], aes(x=datetime, y=air_temp, group = id, colour = id)) + 
    geom_line() +labs(y= "Air temperature loggers", x = "Date") +
    theme(legend.key.size = unit(2,"line"))

  
  
```



# Plot air, inner and outer temp for full id
In general inner and outer show buffered variation compared with air temp EXCEPT for certain times - where much higher peaks - due to direct radiation on trunks?
Very slight phase shift when trunk temps bufferd ()

be_nsp_5 survey tag recoridng about 5C lower than all others
be_oak_1  survey tag show erratic behaviour - strings of 0s and NAs
be_oak_3 - erratic - some 0.000 readings just before string of NAs

NSP6 : very different n-facing trunk temperatures

```{r plot_timeseries}
# Plot temp timeseries
for(id in unique(allresults_df$id)){
  #print(fullid)
  plotdata<-allresults_df[which(allresults_df$id ==id),]
  plot_tseries(plotdata,title=id,legloc='topleft',cols=c('black','cyan','blue','red','orange'))
}
```




```{r plot_part_timeseries}
# Short time series
startdt<-as.POSIXct(strptime("2024-07-22 00:00:00","%Y-%m-%d %H:%M:%S"), tz='GMT')
enddt<-as.POSIXct(strptime("2024-08-06 00:00:00","%Y-%m-%d %H:%M:%S"), tz='GMT')

for(id in unique(allresults_df$id)){
  #print(fullid)
  plotdata<-allresults_df[which(allresults_df$id ==id),]
  startrw<-which(plotdata$datetime==startdt)
  endrw<-which(plotdata$datetime==enddt)
  if(length(startrw)>0 & length(endrw)>0 ) plot_tseries(plotdata[c(startrw:endrw),],title=id,legloc='topleft',cols=c('black','cyan','blue','red','orange'))
}
```

# Plot differences 
Between trunk and air temperatures

Compare difference by time of day - 

```{r plot_difs}
for(id in unique(allresults_df$id)){
  plotdata<-allresults_df[which(allresults_df$id ==id),]
  #plotdata$air_temp[which(plotdata$air_temp==0.00000)]<-NA
  plotdata$si_dif_temp<-plotdata$s_inner_temp-plotdata$air_temp
  plotdata$so_dif_temp<-plotdata$s_outer_temp-plotdata$air_temp
  plotdata$ni_dif_temp<-plotdata$n_inner_temp-plotdata$air_temp
  plotdata$no_dif_temp<-plotdata$n_outer_temp-plotdata$air_temp

  #plotdata$inairdif_temp[which(is.na(plotdata$air_temp))]<-NA
  #plotdata$outairdif_temp[which(is.na(plotdata$air_temp))]<-NA
  
  plotdata<-plotdata[,!(names(plotdata) %in% c("s_inner_temp","s_outer_temp","n_inner_temp","n_outer_temp","air_temp"))]
  plot_tseries(plotdata,title=paste('Temp dif',fullid))
  
  print(paste(fullid,"- max OUTER difference =",round(max(plotdata$outairdif_temp,na.rm=T),2)))
  print(paste(fullid,"- max INNER difference =",round(max(plotdata$inairdif_temp,na.rm=T),2)))
  print(paste(fullid,"- min OUTER difference =",round(min(plotdata$outairdif_temp,na.rm=T),2)))
  print(paste(fullid,"- min INNER difference =",round(min(plotdata$inairdif_temp,na.rm=T),2)))

}




```
## Distribution of temperatures

!!!!  CORRECT  !!!!

```{r distrib_plots}
library(ggplot2)
id<-"cw_nsp_3"

# Combine N and S temp to singel database
treedata<-allresults_df[which(allresults_df$id==id),]




# Only keep observations where data for ALL temperatures (air and trunk)
sel<-which( treedata$s_inner_temp==NA|treedata$s_outer_temp==NA||treedata$n_inner_temp==NA||treedata$n_outer_temp==NA||treedata$air_temp==NA)


ggplot(plotdata.df, aes_(x=as.name(var),col=as.name('group'),fill=as.name('group'))) +
  #geom_histogram(aes(y=..density..), position="identity", alpha=0.3, binwidth = 100) +
  geom_density(alpha=alpha) +
  labs(x=label,y='Density')+
  scale_x_continuous(expand = c(0,0),limits=c(0, maxval))     + 
  scale_y_continuous(expand = c(0,0)) +
  scale_color_manual(values=cols)+
  scale_fill_manual(values=cols)+
  geom_vline(data=mu, aes(xintercept=x, color=Group.1),linetype="dashed")+
  pubtheme() +
  theme(legend.title=element_blank())
}



```
## GDD comparisons
Use treedata to calculate GDD using different methods
Daily Tmx-Tmin / 2
Half hourly T/48
```{r gdd_calcs}

calc_gdd<-function(t,tbase,tdiv){
  dd<-t-tbase
  dd<-ifelse(dd<0,0,dd)
  dd<-dd/tdiv
  gdd<-cumsum(dd)
  return(gdd)
}

calc_daydd<-function(df){
  df$doy<-as.numeric(strftime(df$datetime,format="%j"))
  varnames<-names(df)
  tempvars<-varnames[str_sub(varnames, start=-4)=='temp']
  results_df<-data.frame()[1:length(unique(df$doy)),]
  results_df$doy<-unique(df$doy)
  for(v in tempvars){
    tmax<-by(df[,v],df$doy,max,na.rm=T)
    tmax<-array2DF(tmax,simplify=TRUE) 
    tmin<-by(df[,v],df$doy,min,na.rm=T) 
    tmin<-array2DF(tmin,simplify=TRUE) 
    t<-(tmax$Value+tmin$Value)/2
    results_df[,v]<-t
  }
  return(results_df)
}

#for(tree in unique(allresults_df$id)){
tree<-"be_nsp_4"

# Combine N and S temp to singel database
sdata<-allresults_df[which(allresults_df$fullid==paste0(tree,"_s")),]
ndata<-allresults_df[which(allresults_df$fullid==paste0(tree,"_n")),]
treedata<-sdata[,c("id","datetime","air_temp","inner_temp", "outer_temp")]
names(treedata)<-c("id","datetime","air_temp","s_inner_temp", "s_outer_temp")
treedata<-cbind(treedata,ndata[,c("inner_temp","outer_temp")])
names(treedata)<-c("id","datetime","air_temp","s_inner_temp", "s_outer_temp","n_inner_temp", "n_outer_temp")


# Input data 
startdt<-as.POSIXct(strptime("2024-04-01 00:00:00","%Y-%m-%d %H:%M:%S"), tz='GMT')
enddt<-as.POSIXct(strptime("2024-06-12 00:00:00","%Y-%m-%d %H:%M:%S"), tz='GMT')
indata<-treedata[which(treedata$datetime>=startdt & treedata$datetime<=enddt),]

# Fill missing values in air_temp
library(zoo)
interp_airtemp<-indata[,c('datetime','air_temp')] %>% read.zoo %>% na.approx %>% fortify.zoo
indata$air_temp<-interp_airtemp$.

tbase<-10
tdiv<-48
gd30_df<-data.frame()[1:nrow(indata),]
gd30_df$datetime<-indata$datetime
gd30_df$air_gdd30_temp<-calc_gdd(indata$air_temp,tbase,tdiv) 
gd30_df$n_inner_gdd30_temp<-calc_gdd(indata$n_inner_temp,tbase,tdiv) 
gd30_df$n_outer_gdd30_temp<-calc_gdd(indata$n_outer_temp,tbase,tdiv) 
gd30_df$s_inner_gdd30_temp<-calc_gdd(indata$s_inner_temp,tbase,tdiv) 
gd30_df$s_outer_gdd30_temp<-calc_gdd(indata$s_outer_temp,tbase,tdiv) 

plot_tseries(gd30_df,title=paste(tree,'- Cumulative DD (tbase=',tbase,'30min tstep)'),legloc="topleft")


# Daily GDD min / max
daily_df<-calc_daydd(indata)
tdiv<-1
gdd_df<-data.frame()[1:nrow(daily_df),]
gdd_df$doy<-daily_df$doy
gdd_df$datetime<-as.Date(gdd_df$doy,origin="2023-12-31")
gdd_df$air_gdd_temp<-calc_gdd(daily_df$air_temp,tbase,tdiv) 
gdd_df$n_inner_gdd_temp<-calc_gdd(daily_df$n_inner_temp,tbase,tdiv) 
gdd_df$n_outer_gdd_temp<-calc_gdd(daily_df$n_outer_temp,tbase,tdiv) 
gdd_df$s_inner_gdd_temp<-calc_gdd(daily_df$s_inner_temp,tbase,tdiv) 
gdd_df$s_outer_gdd_temp<-calc_gdd(daily_df$s_outer_temp,tbase,tdiv) 
head(gdd_df)


plot_tseries(gdd_df,title=paste(tree,' - Cumulative GDD (tbase=',tbase,'daily tstep)'),legloc="topleft")


startdt<-as.POSIXct(strptime("2024-05-09 00:00:00","%Y-%m-%d %H:%M:%S"), tz='GMT')
enddt<-as.POSIXct(strptime("2024-05-20 00:00:00","%Y-%m-%d %H:%M:%S"), tz='GMT')

startrw<-which(treedata$datetime==startdt)
endrw<-which(treedata$datetime==enddt)
if(length(startrw)>0 & length(endrw)>0 ) plot_tseries(treedata[startrw:endrw,c("datetime","n_inner_temp","n_outer_temp","air_temp","s_inner_temp","s_outer_temp")],title=tree,legloc='bottomright')

plot_tseries(treedata[startrw:endrw,c("datetime","n_outer_temp","s_outer_temp","air_temp")],title=tree,legloc='bottomright')

plot_tseries(treedata[startrw:endrw,c("datetime","s_inner_temp","s_outer_temp","air_temp")],title=tree,legloc='bottomright')


```

## Apply pest models to timeseries
Compare Dendroctonus (prefers cooler temperatures to Ips

```{r pest_dendroctonus}
id<-"cw_nsp_3"
treedata_df<-allresults_df[which(allresults_df$id==id),]
# remove NAs
treedata_df<-treedata_df[which(complete.cases(treedata_df)),]

Tair<-treedata_df$air_temp
tme<-treedata_df$datetime

columns<-c("pest","site","tree","spp","location","aspect","depth","lay", "hatch", "pupate","adult")
dendro_df<-data.frame(matrix(nrow = 0, ncol = length(columns))) 
colnames(dendro_df) <- columns

# Run for each locataion
out<-"doy"
dendro_df[nrow(dendro_df)+1,]<-c("Dmicans",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","s","outer",
                        run_dendroctonus(treedata_df$s_outer_temp, Tair,tme,out) )

dendro_df[nrow(dendro_df)+1,]<-c("Dmicans",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","s","inner",
                        run_dendroctonus(treedata_df$s_inner_temp, Tair,tme,out) )

dendro_df[nrow(dendro_df)+1,]<-c("Dmicans",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","n","outer",
                        run_dendroctonus(treedata_df$n_outer_temp, Tair,tme,out) )

dendro_df[nrow(dendro_df)+1,]<-c("Dmicans",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","n","inner",
                        run_dendroctonus(treedata_df$n_inner_temp, Tair,tme,out) )

dendro_df[nrow(dendro_df)+1,]<-c("Dmicans",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"air","air","air",
                        run_dendroctonus(treedata_df$air_temp, Tair,tme,out) )
dendro_df
```
```{r pest_ips}
id<-"cw_nsp_3"
treedata_df<-allresults_df[which(allresults_df$id==id),]
# remove NAs
treedata_df<-treedata_df[which(complete.cases(treedata_df)),]

Tair<-treedata_df$air_temp
tme<-treedata_df$datetime

columns<-c("pest","site","tree","spp","location","aspect","depth","emerge", "lay","adult")
ips_df<-data.frame(matrix(nrow = 0, ncol = length(columns))) 
colnames(ips_df) <- columns

# Run for each locataion
out<-"doy" # out<-out
ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","s","outer",
                        run_ips(treedata_df$s_outer_temp, Tair,tme,out) )

ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","s","inner",
                        run_ips(treedata_df$s_inner_temp, Tair,tme,out) )

ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","n","outer",
                        run_ips(treedata_df$n_outer_temp, Tair,tme,out) )

ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","n","inner",
                        run_ips(treedata_df$n_inner_temp, Tair,tme,out) )

ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"air","air","air",
                        run_ips(treedata_df$air_temp, Tair,tme,out) )
ips_df


## 2nd generation 
out<-"doy"

# Run for each locataion
start<-which(treedata_df$datetime==as.POSIXct((as.numeric(ips_df$adult[1])+1)*24*60*60,origin="2024-01-01",tz="GMT"))
             
Tair<-treedata_df$air_temp[start:nrow(treedata_df)]
tme<-treedata_df$datetime[start:nrow(treedata_df)]
T<-treedata_df$s_outer_temp[start:nrow(treedata_df)]
ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","s","outer",
                        run_ips(T, Tair,tme,out) )

start<-which(treedata_df$datetime==as.POSIXct((as.numeric(ips_df$adult[2])+1)*24*60*60,origin="2024-01-01",tz="GMT"))
Tair<-treedata_df$air_temp[start:nrow(treedata_df)]
tme<-treedata_df$datetime[start:nrow(treedata_df)]
T<-treedata_df$s_inner_temp[start:nrow(treedata_df)]
ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","s","inner",
                        run_ips(T, Tair,tme,out) )

start<-which(treedata_df$datetime==as.POSIXct((as.numeric(ips_df$adult[3])+1)*24*60*60,origin="2024-01-01",tz="GMT"))
Tair<-treedata_df$air_temp[start:nrow(treedata_df)]
tme<-treedata_df$datetime[start:nrow(treedata_df)]
T<-treedata_df$n_outer_temp[start:nrow(treedata_df)]
ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","n","outer",
                        run_ips(T, Tair,tme,out) )

start<-which(treedata_df$datetime==as.POSIXct((as.numeric(ips_df$adult[4])+1)*24*60*60,origin="2024-01-01",tz="GMT"))
Tair<-treedata_df$air_temp[start:nrow(treedata_df)]
tme<-treedata_df$datetime[start:nrow(treedata_df)]
T<-treedata_df$n_inner_temp[start:nrow(treedata_df)]
ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"bark","n","inner",
                        run_ips(T, Tair,tme,out) )

start<-which(treedata_df$datetime==as.POSIXct((as.numeric(ips_df$adult[5])+1)*24*60*60,origin="2024-01-01",tz="GMT"))
Tair<-treedata_df$air_temp[start:nrow(treedata_df)]
tme<-treedata_df$datetime[start:nrow(treedata_df)]
ips_df[nrow(ips_df)+1,]<-c("Ipstypog",treedata_df$site[1],
                         treedata_df$tree[1],treedata_df$spp[1],"air","air","air",
                        run_ips(Tair, Tair,tme,out) )
ips_df

ips_df$id<-paste(ips_df$location,ips_df$aspect,ips_df$depth, ips_df$generation, sep="_")
grp1<-which(ips_d)
t(ips_df[which(ips_df$id=="bark_s_outer_1"),c(8,11)])
```
